---
title: "England Blanket Bog Model: Create training data"
output: html_notebook
---

## Packages 
```{r}
library(raster)
library(rgdal)

library(tidyverse)
```

## Import data

### Import observations: land classification file 
```{r}
classes <- "../data/FEP_points/Catchment_FEP_OS_TrainingPoints_final_clean.shp"
ogrInfo(classes)

#save the shapefile into an object
observationsSPDF <- readOGR(dsn = classes, verbose = TRUE)

rm(classes)
```

### Import predictor data
```{r}
# Import environmental and topographic data
#load(file = "../data/topo_env_data.rda")

elev <- raster("../data/topo_env_data.tif", band = 1)
surf <- raster("../data/topo_env_data.tif", band = 2)
inflow <- raster("../data/topo_env_data.tif", band = 3)
outflow <- raster("../data/topo_env_data.tif", band = 4)
slope <- raster("../data/topo_env_data.tif", band = 5)
aspect <- raster("../data/topo_env_data.tif", band = 6)

# Import UKCP09 climate data
#load(file = "../data/ukcp09GriddedDataAnnPrepd.rda")
gdd_1960_90_ann <- raster("../data/ukcp09GriddedDataAnnPrepd.tif", band = 1)
gsl_1960_90_ann <- raster("../data/ukcp09GriddedDataAnnPrepd.tif", band = 2)
rain_1960_90_annual <- raster("../data/ukcp09GriddedDataAnnPrepd.tif", band = 3)
rain_daily_mean_1960_90_annual <- raster("../data/ukcp09GriddedDataAnnPrepd.tif", band = 4)
raindays_1mm_1960_90_annual <- raster("../data/ukcp09GriddedDataAnnPrepd.tif", band = 5)
raindays_10mm_1960_90_annual <- raster("../data/ukcp09GriddedDataAnnPrepd.tif", band = 6)
temp_mean_1960_90_annual <- raster("../data/ukcp09GriddedDataAnnPrepd.tif", band = 7)
temp_min_1960_90_annual <- raster("../data/ukcp09GriddedDataAnnPrepd.tif", band = 8)
temp_max_1960_90_annual <- raster("../data/ukcp09GriddedDataAnnPrepd.tif", band = 9)

```

## prepare observations
```{r}
#raster::extract metadata
boundbox <- observationsSPDF@bbox; boundbox 
CRS <- observationsSPDF@proj4string
```

```{r}
#filter observations
observations <- observationsSPDF[which(observationsSPDF@data$Tier == 1 | observationsSPDF@data$Tier == 2),]
summary(observations); class(observations)
unique(observations$Detailed)
unique(observations$Broad)

rm(observationsSPDF)
```

```{r}
# make a logical vector for blanket bog presence
observations$bbog <- if_else(observations$Detailed == "Blanket Bog", true = "present", false = "absent")
```

reduce the number of absences to balance the dataset
```{r}
indexSubset <- c(
  sample(x = which(observations$bbog %in% "absent"), #a subset of absences
       size = which(observations$bbog %in% "present"), #same number as presence
       replace = FALSE), 
  which(observations$bbog %in% "present") #index of presence
)
length(indexSubset); str(indexSubset)

observations <- observations[indexSubset,]; dim(observations) #subset the observations dataset
```


## Create input dataset

```{r}
#create dataframe with peat depth as first col
input.data <- data.frame(bbog = observations$bbog)

#raster::extract environmental and topographic data for each peat depth measurement
input.data$elev <- raster::extract(
  elev, coordinates(observations)[,1:2])
input.data$aspect <- raster::extract(
  aspect, coordinates(observations)[,1:2])
input.data$slope <- raster::extract(
  slope, coordinates(observations)[,1:2])
input.data$outflow <- raster::extract(
  outflow, coordinates(observations)[,1:2])
input.data$inflow <- raster::extract(
  inflow, coordinates(observations)[,1:2])
input.data$surf <- raster::extract(
  surf, coordinates(observations)[,1:2])

#raster::extract climate data for each peat depth measurement
input.data$gdd <- raster::extract(
  gdd_1960_90_ann, coordinates(observations)[,1:2])
input.data$gsl <- raster::extract(
  gsl_1960_90_ann, coordinates(observations)[,1:2])
input.data$rain_ann <- raster::extract(
  rain_1960_90_annual, coordinates(observations)[,1:2])
input.data$rain_daily <- raster::extract(
  rain_daily_mean_1960_90_annual, coordinates(observations)[,1:2])
input.data$raindays_10mm <- raster::extract(
  raindays_10mm_1960_90_annual, coordinates(observations)[,1:2])
input.data$raindays_1mm <- raster::extract(
  raindays_1mm_1960_90_annual, coordinates(observations)[,1:2])
input.data$temp_mean <- raster::extract(
  temp_mean_1960_90_annual, coordinates(observations)[,1:2])
input.data$temp_min <- raster::extract(
  temp_min_1960_90_annual, coordinates(observations)[,1:2])
input.data$temp_max <- raster::extract(
  temp_max_1960_90_annual, coordinates(observations)[,1:2])
```

### remove errors and outliers from training data
```{r}
summary(input.data)
```

looking at the data frame and the summary above, there are a few errors.  There are 600 records where the elevation is -1000, all of these are not blanket bog.  Also there are 24 records where the climate data is NA (these coincide with the -1000 elevation.)  We will remove the records where elevation is -1,000. 

```{r}
input.data <- filter(input.data, elev != -1000)
summary(input.data)
```






## export to file

```{r}
save(input.data, file="../data/input.data.rds")
write.csv(input.data, file = "../data/input_data.csv", row.names = F)
```
