---
title: "England Peat Depth Model: Train models"
output: html_notebook
---

#Peat Depths: run models

## Packages 
```{r, include=FALSE}
library(raster)
library(rgdal)

library(randomForest)
library(Rborist)
library(e1071)
library(kernlab)

library(caret)
library(parallel)
library(doParallel)

library(aws.ec2metadata)

```

## Load data
```{r}
load(file = "../data/input.data.rds")
dim(input.data)
```

### create directory if necessary
```{r}
dir.create("../data/models")
```


## Start Parallel Processing
```{r}
detectCores()
getDoParWorkers()
cl <- makeCluster(detectCores()-1)
registerDoParallel(cl)
getDoParName()
getDoParWorkers()
#registerDoSEQ() # to stop parallel processing
```


## Make training and test sets

```{r}
set.seed(123)
inTrain <- createDataPartition(y = input.data$PEAT_DEPTH, 
                               p = .75, # The percentage of data in the training set
                               list = FALSE) # The format of the results

training <- input.data[inTrain,]
testing <- input.data[-inTrain,]
```
Training data has `r nrow(training)` records.  Testing data has `r nrow(testing)` records.  

#### train model
**Need the latest version of caret from github to avoid known bug.**

set model parameters
```{r}
params <- trainControl(method = "cv", number = 10)
```

set up system performance monitoring
```{r}
modeltime <- data.frame(model = as.character(), 
                        insType = as.character(),
                        time_mins = as.difftime(character()))
```



train linear model
```{r}
  starttime <- Sys.time()

Mlmuk1a <- train(PEAT_DEPTH ~ ., 
              data = training, 
              method = "lm")

  modeltime <- rbind(modeltime, 
                     data.frame(model = "Mlmuk1a",
                                insType = aws.ec2metadata::metadata$instance_type(),
                                time_mins = difftime(Sys.time(), starttime, units = "mins")))

save(Mlmuk1a, file = "../data/models/Mlmuk1a.rds")

Mlmuk1a
```

train random forest model
```{r}
  starttime <- Sys.time()

Mrfuk1a <- train(PEAT_DEPTH ~ ., 
              data = training, 
              method = "Rborist")

  modeltime <- rbind(modeltime, 
                     data.frame(model = "Mrfuk1a",
                                insType = aws.ec2metadata::metadata$instance_type(),
                                time_mins = difftime(Sys.time(), starttime, units = "mins")))  
save(Mrfuk1a, file = "../data/models/Mrfuk1a.rds")

Mrfuk1a
```


```{r}
rbind(getTrainPerf(Mlmuk1a), getTrainPerf(Mrfuk1a))
```

train generalised linear model
```{r}
  starttime <- Sys.time()

Mglmuk1a <- train(PEAT_DEPTH ~ ., 
              data = training, 
              method = "glm")

  modeltime <- rbind(modeltime, 
                     data.frame(model = "Mglmuk1a",
                                insType = aws.ec2metadata::metadata$instance_type(),
                                time_mins = difftime(Sys.time(), starttime, units = "mins")))
save(Mglmuk1a, file = "../data/models/Mglmuk1a.rds")

Mglmuk1a
```

train boosted generalised linear model
```{r}
  starttime <- Sys.time()

Mbglmuk1a <- train(PEAT_DEPTH ~ ., 
              data = training, 
              method = "glmboost")


  modeltime <- rbind(modeltime, 
                     data.frame(model = "Mbglmuk1a",
                                insType = aws.ec2metadata::metadata$instance_type(),
                                time_mins = difftime(Sys.time(), starttime, units = "mins")))
save(Mbglmuk1a, file = "../data/models/Mbglmuk1a.rds")

Mbglmuk1a
```

train k nearest neighbours model

```{r}
  starttime <- Sys.time()

Mknnuk1a <- train(PEAT_DEPTH ~ ., 
              data = training, 
              method = "knn")

  modeltime <- rbind(modeltime, 
                     data.frame(model = "Mknnuk1a",
                                insType = aws.ec2metadata::metadata$instance_type(),
                                time_mins = difftime(Sys.time(), starttime, units = "mins")))
  
save(Mknnuk1a, file = "../data/models/Mknnuk1a.rds")

Mknnuk1a
```


train Generalised Additive Model 

```{r}
  starttime <- Sys.time()

Mgamuk1a <- train(PEAT_DEPTH ~ ., 
              data = training, 
              method = "gam")

  modeltime <- rbind(modeltime, 
                     data.frame(model = "Mgamuk1a",
                                insType = aws.ec2metadata::metadata$instance_type(),
                                time_mins = difftime(Sys.time(), starttime, units = "mins")))
  
save(Mgamuk1a, file = "../data/models/Mgamuk1a.rds")

Mgamuk1a
```

train Gradient Boosting Model 

```{r}
  starttime <- Sys.time()

Mgbmuk1a <- train(PEAT_DEPTH ~ ., 
              data = training, 
              method = "gbm")

  modeltime <- rbind(modeltime, 
                     data.frame(model = "Mgbmuk1a",
                                insType = aws.ec2metadata::metadata$instance_type(),
                                time_mins = difftime(Sys.time(), starttime, units = "mins")))
  
save(Mgbmuk1a, file = "../data/models/Mgbmuk1a.rds")

Mgbmuk1a
```

train support vector machine (SVM)

```{r}
#standardise training data first
#preproc_stand <- preProcess(training, method = c("center", "scale"))
#training_stand <- predict(preproc_stand, training); summary(training_stand)
  starttime <- Sys.time()

Msvmuk1a <- train(PEAT_DEPTH ~ ., 
                data = training, 
                method = "svmLinear", # Linear Kernel
                preProc = c("center", "scale")) #standardise data

  modeltime <- rbind(modeltime, 
                     data.frame(model = "Msvmuk1a",
                                insType = aws.ec2metadata::metadata$instance_type(),
                                time_mins = difftime(Sys.time(), starttime, units = "mins")))
  
save(Msvmuk1a, file = "../data/models/Msvmuk1a.rds")

Msvmuk1a
```


## Evaluate models
```{r}
#load models
load("../data/models/Mlmuk1a.rds")
load("../data/models/Mrfuk1a.rds")
load("../data/models/Mglmuk1a.rds")
load("../data/models/Mbglmuk1a.rds")
load("../data/models/Mgamuk1a.rds")
load("../data/models/Mgbmuk1a.rds")
load("../data/models/Msvmuk1a.rds")
```

```{r}
cvValues <- resamples(list(LM = Mlmuk1a, RF = Mrfuk1a, GLM = Mglmuk1a, GLMB = Mbglmuk1a, 
                           GAM = Mgamuk1a, GBM = Mgbmuk1a, SVM = Msvmuk1a))
summary(cvValues)
splom(cvValues, metric = "Rsquared")
parallelplot(cvValues, metric = "Rsquared")
dotplot(cvValues, metric = "Rsquared")
```

```{r}
Diffs <- diff(cvValues, metric = "Rsquared")
summary(Diffs)
dotplot(Diffs, metric = "Rsquared")
```

## predict test data from model and evaluate

```{r}
#transform testing data

preproc_stand <- preProcess(testing, method = c("center", "scale"))
testing_stand <- predict(preproc_stand, testing)
```


```{r}
lm_pred <- predict(Mlmuk1a, testing)

rf_pred <- predict(Mrfuk1a, testing)

glm_pred <- predict(Mglmuk1a, testing)

glmb_pred <- predict(Mbglmuk1a, testing) 

gam_pred <- predict(Mgamuk1a, testing)

gbm_pred <- predict(Mgbmuk1a, testing)

svm_pred <- predict(Msvmuk1a, testing)

```

Calculate deviation

**RMSE**: Root Mean Square Error. Standard deviation of the differences between predicted and observed values.  Measures accuracy.  Scale-dependent, so can't compare different datasets.  Sensitive to outliers. Smaller is better.  Calculated by   `sqrt(mean((pred - obs)^2` 

**Rsquared**: Coefficient of determination.  The proportion of the variation in the observed variable that is predictable from the predicted variable.  Not scale dependent.  Closer to 1 is better. Calculated by `R^2 = 1-\frac{∑ (y_i - \hat{y}_i)^2}{∑ (y_i - \bar{y}_i)^2}`

**MAE**: Mean Absolute Error.  The average of the difference between the absolute predicted and observed values.  Scale-dependent, so can't compare different datasets.  Sensitive to outliers. Smaller is better.  Calculated by `mean(abs(pred - obs))`


```{r}
mod_variance <- data.frame(rbind(
  postResample(pred = lm_pred, obs = testing$PEAT_DEPTH),
  postResample(pred = rf_pred, obs = testing$PEAT_DEPTH),
  postResample(pred = glm_pred, obs = testing$PEAT_DEPTH),
  postResample(pred = glmb_pred, obs = testing$PEAT_DEPTH),
  postResample(pred = gam_pred, obs = testing$PEAT_DEPTH),
  postResample(pred = gbm_pred, obs = testing$PEAT_DEPTH),
  postResample(pred = svm_pred, obs = testing$PEAT_DEPTH)
))

mod_variance <- cbind(model = c("lm", "rf", "glm", "glmb", "gam", "gbm", "svm"), mod_variance)
mod_variance
```




## Train full models

```{r}
  starttime <- Sys.time()

Mlmuk1a.full <- train(PEAT_DEPTH ~ ., 
              data = input.data, 
              method = "lm")

  modeltime <- rbind(modeltime, 
                     data.frame(model = "Mlmuk1a.full",
                                insType = aws.ec2metadata::metadata$instance_type(),
                                time_mins = difftime(Sys.time(), starttime, units = "mins")))
  
save(Mlmuk1a.full, file = "../data/models/Mlmuk1a.full.rds")

Mlmuk1a.full
```

train model
```{r}
  starttime <- Sys.time()

Mrfuk1a.full <- train(PEAT_DEPTH ~ ., 
              data = input.data, 
              method = "Rborist")

  modeltime <- rbind(modeltime, 
                     data.frame(model = "Mrfuk1a.full",
                                insType = aws.ec2metadata::metadata$instance_type(),
                                time_mins = difftime(Sys.time(), starttime, units = "mins")))
  
save(Mrfuk1a.full, file = "../data/models/Mrfuk1a.full.rds")

Mrfuk1a.full
```

evaluate
```{r}
rbind(getTrainPerf(Mlmuk1a.full), getTrainPerf(Mrfuk1a.full))
```

