---
title: 'England Peat Depth Model: Data Preparation: Predictors'
output:
  html_document: default
  html_notebook: default
---


## Packages 
```{r, include=FALSE}
library(raster)
library(rgdal)
library(tidyverse)
```


## Import data ----
```{r}
list.files(path = "../data")
```



## Import environmental and topographic data
```{r}
#load(file = "../data/topo_env_data_2.rda")
#elev <- read_rds(path = "../data/elev.rds")
load("../data/elev.Rdata")
load("../data/surf.Rdata")
load("../data/predictor_stack.Rdata")
plot(elev)
plot(predictors$elev)
```

### Calculate derrived data
 
```{r}
slope <- raster::terrain(x = elev, opt='slope', unit='radians', neighbors=8)

aspect <- terrain(x = elev, opt = "aspect", unit = "degrees",
                  neighbors = 8, filename = "../data/slope.tif")
```




## Save environmental and topographic data to file
```{r}
#save to raster stack then geotiff (might work better than .rda)
topo_env_stack <- stack(elev, surf, inflow, outflow, slope, aspect, moorline)
writeRaster(x = topo_env_stack, filename = "../data/topo_env_data.tif", overwrite = TRUE)

#save(elev, surf, inflow, outflow, slope, aspect, file = "../data/topo_env_data.rda")
```

## Import UKCP09 climate data

These data are Met Office figures produced for the UK Climate Projections 2009 and available from www.ceda.ac.uk (collection ukcp09, 5km gridded long term averages).  The process for reading and converting them to raster format, as well as the calculation of derived values (such as seasonal figures) are contained in the script `ukcp09DataImport.R`.  

```{r}
# load data rasters

# Annual data rasters
load(file = "../data/ukcp09GriddedDataAnn.rda")

# Seasonal data rasters
# load(file = "../data/ukcp09GriddedDataSease.rda")

# monthly data rasters - lots of objects
# load(file = "../data/ukcp09GriddedDataMonthly.rda")
```

resample rasters
```{r}
#resample: x = raster to be resampled, y = raster to be resampled to
ls(pattern = "1960_90")
gdd_1960_90_ann <- resample(gdd_1960_90_ann, elev)
gsl_1960_90_ann <- resample(gsl_1960_90_ann, elev)
rain_1960_90_annual <- resample(rain_1960_90_annual, elev)
rain_daily_mean_1960_90_annual <- resample(rain_daily_mean_1960_90_annual, elev)
raindays_10mm_1960_90_annual <- resample(raindays_10mm_1960_90_annual, elev)
raindays_1mm_1960_90_annual <- resample(raindays_1mm_1960_90_annual, elev)
temp_max_1960_90_annual <- resample(temp_max_1960_90_annual, elev)
temp_mean_1960_90_annual <- resample(temp_mean_1960_90_annual, elev)
temp_min_1960_90_annual <- resample(temp_min_1960_90_annual, elev)
```

### Save resampled climate data
```{r}
climate_stack <- stack(gdd_1960_90_ann, gsl_1960_90_ann,
  rain_1960_90_annual, rain_daily_mean_1960_90_annual,
  raindays_1mm_1960_90_annual, raindays_10mm_1960_90_annual,
  temp_mean_1960_90_annual, temp_min_1960_90_annual, 
  temp_max_1960_90_annual)
writeRaster(x = climate_stack, filename = "../data/ukcp09GriddedDataAnnPrepd.tif")
rm(climate_stack)
```



