---
title: "England Peat Depth Model: Data Preparation"
output: html_notebook
---

#Peat Depths: data preparation script

## Packages 
```{r}
library(raster)
library(rgdal)
library(eply) # for unquoting ls objects
```


## Import data ----
```{r}
list.files(path = "../data")
list.files(path = "../data/peat_depth_data/")
```

### Import the main file geodatabase 

I don't think we need this, we're basically recreating it.  Its the 'input data' from the old model.  currently peat depth data is not in there
```{r}
# fgdb <- "../data/PeatDepth.gdb"
#   # List all feature classes in a file geodatabase
#   subset(ogrDrivers(), grepl("GDB", name))
#   fc_list <- ogrListLayers(fgdb)
#   print(fc_list)
```



## Import environmental and topographic data
```{r}
# #elevation 
# elev <- raster("../data/Ground Elevation © CEH1.tif")
# 
# #surface type 
# surf <- raster("../data/Surface Type © CEH1.tif")
# 
# #inflow pattern
# inflow <- raster("../data/Inflow Pattern © CEH1.tif")
# 
# #outflow pattern
# outflow <- raster("../data/Outflow Pattern © CEH1.tif")
#
```

### Calculate derrived data
 
```{r}
# slope <- terrain(x = elev, opt = "slope", unit = "radians", neighbors = 8)
# 
# aspect <- terrain(x = elev, opt = "aspect", unit = "degrees", neighbors = 8)
```

## Save environmental and topographic data to file
```{r}
# #save to file
# save(list = c("elev", "surf", "inflow", "outflow", "slope", "aspect"), file = "../data/topo_env_data.rda")
```

## Import UKCP09 climate data

These data are Met Office figures produced for the UK Climate Projections 2009 and available from www.ceda.ac.uk (collection ukcp09, 5km gridded long term averages).  The process for reading and converting them to raster format, as well as the calculation of derived values (such as seasonal figures) are contained in the script `ukcp09DataImport.R`.  

```{r}
# load data rasters

# Annual data rasters
load(file = "../data/ukcp09GriddedDataAnn.rda")

# Seasonal data rasters
# load(file = "../data/ukcp09GriddedDataSease.rda")

# monthly data rasters - lots of objects
# load(file = "../data/ukcp09GriddedDataMonthly.rda")
```

resample rasters
```{r}
#resample: x = raster to be resampled, y = raster to be resampled to
ls(pattern = "1960_90")
gdd_1960_90_ann <- resample(gdd_1960_90_ann, elev)
gsl_1960_90_ann <- resample(gsl_1960_90_ann, elev)
rain_1960_90_annual <- resample(rain_1960_90_annual, elev)
rain_daily_mean_1960_90_annual <- resample(rain_daily_mean_1960_90_annual, elev)
raindays_10mm_1960_90_annual <- resample(raindays_10mm_1960_90_annual, elev)
raindays_1mm_1960_90_annual <- resample(raindays_1mm_1960_90_annual, elev)
temp_max_1960_90_annual <- resample(temp_max_1960_90_annual, elev)
temp_mean_1960_90_annual <- resample(temp_mean_1960_90_annual, elev)
temp_min_1960_90_annual <- resample(temp_min_1960_90_annual, elev)
```

### Save resampled climate data
```{r}
# save all annual grids ----
  save(file = "../data/ukcp09GriddedDataAnnPrepd.rds", list = c(
    "gdd_1960_90_ann", "gsl_1960_90_ann",
    "rain_1960_90_annual", "rain_daily_mean_1960_90_annual",
    "raindays_1mm_1960_90_annual", "raindays_10mm_1960_90_annual",
    "temp_mean_1960_90_annual", "temp_min_1960_90_annual", 
    "temp_max_1960_90_annual"))
```


That's it, now go to peat_depth_make_training_data.Rmd